Quantifying Infants' Everyday Restrained Experiences in the Home Using Wearable Inertial Sensors

### 2025-11-09 manuscript_restrained_classification validation.qmd
Fixed typo

author-note:
  blank-lines-above-author-note: 1
  # Disclosures condensed to one paragraph, but you can start a field with two line breaks to break them up: \n\nNew Paragraph
  disclosures:
    # Example: This study was registered at ClinicalTrials.gov (Identifier NTC998877).
    study-registration: null
    # Acknowledge and cite data/materials to be shared.
    data-sharing: null
    # Example: This article is based on data published in Pulaski (2017).
    # Example: This article is based on the dissertation completed by Graham (2018).    
    # Example: Sally Jones earns royalties from the sale of Test X.
    conflict-of-interest: null
    # Example: This study was supported by Grant A123 from the National Science Foundation.
    financial-support: null
    # Example: The authors are grateful for the technical assistance of Dr. X during the initial design and setup of our lab equipment.
    gratitude: null
    # Example. Because the authors are equal contributors, order of authorship was determined by a fair coin toss.
    authorship-agreements: null
abstract: "Physical restraint—including being held, carried, and restrained in devices—is a common feature of infants’ everyday lives. This study developed and validated a machine learning model to detect infants’ restraint using wearable sensors movement data derived from an existing data set (Franchak, 2023). The model achieved high accuracy (89%) and substantial kappa agreement (0.73) compared with human-coded ground truth. The model presented a slight bias towards overestimating unrestrained periods compared to restrained periods, but the bias was mitigated when using a longer data-aggregating window length. The model also showed convergent validity by corroborating prior studies that showed an age-related decrease in infant restraint time. In short, the current study demonstrated the feasibility of using wearable sensors to quantify infants’ real-world restraint experiences, offering a new tool for studying how daily restraint influences early development."
keywords: [infant, restraint, everyday experiences, wearable sensors, machine learning]
bibliography: [references.bib]
format:
  apaquarto-docx: 
    floatsintext: false
  apaquarto-html: default
  apaquarto-typst: default
  apaquarto-pdf: 
    # can be jou (journal), man (manuscript), stu (student), or doc (document)
    # for now, tables and figures do not render properly in jou mode. 
    documentmode: man
    # can be 10pt, 11pt, or 12pt
    fontsize: 12pt
    # Integrate tables and figures in text
    floatsintext: false
    # a4 paper if true, letter size if false
    a4paper: false
    # suppresses loading of the lmodern font package
    nolmodern: false
    # Suppresses loading of the fontenc package
    nofontenc: false
    # Suppress the title above the introduction
    donotrepeattitle: false
    # In jou mode, use times or pslatex instead of txfonts
    notxfonts: false
    # In jou mode, use Computer Modern font instead of times
    notimes: false
    # In jou mode, cancels automatic stretching of tables to column width 
    notab: false
    # Uses Helvetica font in stu and man mode
    helv: false
    # In man and stu mode, neutralizes the \helvetica command
    nosf: false
    # In man and stu mode, uses typewriter font
    tt: false
    # Puts draft watermark on first page
    draftfirst: false
    # Puts draft watermark on all pages
    draftall: false
    # Masks references that are marked as the author's own
    mask: false
    journal: null
    volume: null
    course: null
    professor: null
    duedate: null
    # Hides correspondence text
    nocorrespondence: false
editor_options: 
  chunk_output_type: console
---

@@

```{r fig-cibtwctg}
#| fig-cap: Differences in sensitivity and PPV between restrained and unrestrained prediction according to window lengths
#| fig-width: 8
#| fig-height: 5
#| apa-note: Each vertical line indicates the 95%CI of the metric's difference between restrained prediction and unrestrained prediction. The dot on each line denotes the mean of the difference (restrained - unrestrained).

@@

The prediction bias was also evident when examining the model’s overall estimate of restrained time and its difference from human annotation. @fig-prevalencegap B shows that the model predictions consistently underestimated overall restrained time compared to human annotation, despite the high correlation between the two. We defined prediction error as the model predicted prevalence (in % units) subtracted by the human annotated prevalence of restrained periods (in % units). The 4 s window had a prediction error of `r apa_num(table_prev[1,5]*100)`%, 95% CI \[`r apa_num(table_prev[1,6]*100)`%, `r apa_num(table_prev[1,7]*100)`%\], the 16 s window of `r apa_num(table_prev[2,5])*100`%, 95% CI \[`r apa_num(table_prev[2,6]*100)`%, `r apa_num(table_prev[2,7]*100)`%\], and the 30 s window of `r apa_num(table_prev[3,5]*100)`%, 95% CI \[`r apa_num(table_prev[3,6]*100)`%, `r apa_num(table_prev[3,7]*100)`%\].

>>>>

author-note:
  blank-lines-above-author-note: 1
  # Disclosures condensed to one paragraph, but you can start a field with two line breaks to break them up: \n\nNew Paragraph
  disclosures:
    # Acknowledge and cite data/materials to be shared.
    # Example. Because the authors are equal contributors, order of authorship was determined by a fair coin toss.
    # Example: This study was registered at ClinicalTrials.gov (Identifier NTC998877).
    study-registration: null
    data-sharing: null
    # Example: This article is based on data published in Pulaski (2017).
    # Example: This article is based on the dissertation completed by Graham (2018).    
    # Example: Sally Jones earns royalties from the sale of Test X.
    conflict-of-interest: null
    # Example: This study was supported by Grant A123 from the National Science Foundation.
    financial-support: null
    # Example: The authors are grateful for the technical assistance of Dr. X during the initial design and setup of our lab equipment.
    gratitude: null
    authorship-agreements: null
abstract: "Physical restraint—including being held, carried, and restrained in devices—is a common feature of infants’ everyday lives. This study developed and validated a machine learning model to detect infants’ restraint using wearable sensors movement data derived from an existing data set (Franchak, 2023). The model achieved high accuracy (89%) and substantial kappa agreement (0.73) compared with human-coded ground truth. The model presented a slight bias towards overestimating unrestrained periods compared to restrained periods, but the bias was mitigated when using a longer data-aggregating window length. The model also showed convergent validity by corroborating prior studies that showed an age-related decrease in infant restraint time. In short, the current study demonstrated the feasibility of using wearable sensors to quantify infants’ real-world restraint experiences, offering a new tool for studying how daily restraint influences early development."
keywords: [infant, restraint, everyday experiences, wearable sensors, machine learning]
bibliography: [references.bib]
format:
  apaquarto-docx: 
    floatsintext: false
  apaquarto-html: default
  apaquarto-typst: default
  apaquarto-pdf: 
    # can be jou (journal), man (manuscript), stu (student), or doc (document)
    # for now, tables and figures do not render properly in jou mode. 
    # can be 10pt, 11pt, or 12pt
    # Integrate tables and figures in text
    # a4 paper if true, letter size if false
    # suppresses loading of the lmodern font package
    # Suppresses loading of the fontenc package
    # Suppress the title above the introduction
    # In jou mode, use times or pslatex instead of txfonts
    # In jou mode, use Computer Modern font instead of times
    # In jou mode, cancels automatic stretching of tables to column width 
    # Uses Helvetica font in stu and man mode
    # In man and stu mode, neutralizes the \helvetica command
    # In man and stu mode, uses typewriter font
    # Puts draft watermark on first page
    # Puts draft watermark on all pages
    # Masks references that are marked as the author's own
    # Hides correspondence text
    documentmode: man
    fontsize: 12pt
    floatsintext: false
    a4paper: false
    nolmodern: false
    nofontenc: false
    donotrepeattitle: false
    notxfonts: false
    notimes: false
    notab: false
    helv: false
    nosf: false
    tt: false
    draftfirst: false
    draftall: false
    mask: false
    journal: null
    volume: null
    course: null
    professor: null
    duedate: null
    nocorrespondence: false
editor_options: 
  chunk_output_type: console
---

@@

```{r fig-cibtwctg}
#| fig-cap: Differences in sensitivity and PPV between restrained and unrestrained prediction according to window length
#| fig-width: 8
#| fig-height: 5
#| apa-note: Each vertical line indicates the 95%CI of the metric's difference between restrained prediction and unrestrained prediction. The dot on each line denotes the mean of the difference (restrained - unrestrained).

@@

The prediction bias was also evident when examining the model’s overall estimate of restrained time and its difference from human annotation. @fig-prevalencegap B shows that the model predictions consistently underestimated overall restrained time compared to human annotation, despite the high correlation between the two. We defined prediction error as the model predicted prevalence (in % units) subtracted by the human annotated prevalence of restrained periods (in % units). The 4 s window had a prediction error of `r apa_num(table_prev[1,5]*100)`%, 95% CI \[`r apa_num(table_prev[1,6]*100)`%, `r apa_num(table_prev[1,7]*100)`%\], the 16 s window of `r apa_num(table_prev[2,5]*100)`%, 95% CI \[`r apa_num(table_prev[2,6]*100)`%, `r apa_num(table_prev[2,7]*100)`%\], and the 30 s window of `r apa_num(table_prev[3,5]*100)`%, 95% CI \[`r apa_num(table_prev[3,6]*100)`%, `r apa_num(table_prev[3,7]*100)`%\].
